Mapping plant-pollinator interactions across Europe

### 2023-12-11 Manuscript/Ms.qmd
map

# Abstract

Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants. However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes. One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time.

@@

# Results

>>>>

```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(giscoR)
library(sf)
```
# Abstract

Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants. However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes. One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time. Here, we present a dataset of plant-pollinator interactions at European level that consists in XXX studies compiled during the years xxx and xxxx.

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
#Print number of countries
data = readRDS("../Data/Interactions_uncounted.rds")
#Country number
country_number = data %>% 
select(Country) %>% 
n_distinct()
#Study number
study_number = data %>% 
select(Study_id) %>% 
n_distinct()
#Minimum year
min_year = data %>% 
select(Year) %>% 
slice(which.min(Year)) %>% 
pull()
#Maximum year
max_year = data %>% 
select(Year) %>% 
slice(which.max(Year)) %>% 
pull()
#Number of interactions
interaction_number = nrow(data)
#Check number of accepted pollinators
pollinator_spp_number = data %>% 
select(Pollinator_rank, Pollinator_accepted_name)%>%
filter(Pollinator_rank == "SPECIES") %>% 
select(Pollinator_accepted_name) %>% 
n_distinct()
#Check number of plant species
plant_spp_number = data %>% 
select(Plant_rank, Plant_status, 
Plant_matchtype, Plant_accepted_name) %>% 
filter(Plant_rank == "SPECIES") %>% 
select(Plant_accepted_name) %>% 
n_distinct()

```

## Dataset description

This European metaweb consist of datasets of plant-pollinator interactions compiled initially by a wide number of researchers and institutions within the European continent. This dataset covers `r country_number` countries and consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year`, and accounts for a total of `r prettyNum(interaction_number, big.mark=",")` interactions from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.

## Species coverage

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="100%"}
#Prepare coordinates
for_plotting = data %>% 
select(Study_id, Latitude, Longitude) %>% 
distinct()
#Get countries and transfrom coordinates
countries <- gisco_get_countries(
resolution = 20) %>%
st_transform(3035) %>% 
sf::st_as_sf(crs = 4326)

#Number of plant and poll species per network
#-----------------------#
#Pollinators------
#-----------------------#
#Check unique number of pollinator species
PollNumber_by_ID = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
select(Pollinator_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
#Merge now with spp counts by network dataset
poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
#Create duplicate cord columns to operate on them
poll_data$coords <- data.frame(
Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
#Set coordinates
poll_data$coords_sf = st_as_sf(poll_data$coords, 
  coords = c("Longitude", "Latitude"), crs = 4326)
#Plot map of Europe
#Very noisy with many coordinates per point,
#Let's select a unique coordinate by study
poll_data1 =  poll_data %>% 
distinct(Study_id, .keep_all = TRUE)
#-----------------------#
#Plants-----
#-----------------------#
#Check unique number of plant species
PlantNumber_by_ID = data %>% 
select(Plant_rank, Plant_accepted_name, Study_id) %>% 
filter(Plant_rank == "SPECIES") %>% 
select(Plant_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
summarise(Plant_counts = n_distinct(Plant_accepted_name))
#Merge now with spp counts by network dataset
plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
#Create duplicate cord columns to operate on them
plant_data$coords <- data.frame(
Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
#Set coordinates
plant_data$coords_sf = st_as_sf(plant_data$coords, 
  coords = c("Longitude", "Latitude"), crs = 4326)
#Plot map of Europe
#Very noisy with many coordinates per point,
#Let's select a unique coordinate by study
plant_data1 =  plant_data %>% 
distinct(Study_id, .keep_all = TRUE)
#Add vars
plant_data1$Group = "Plant"
poll_data1$Group = "Pollinator"
#Rename cols
plant_data2 = plant_data1 %>% 
rename(Counts = Plant_counts)
poll_data2 = poll_data1 %>% 
rename(Counts = Pollinator_counts)
#Bind rows
all_data = bind_rows(plant_data2, poll_data2)

#Try to plot both together (same plot)
ggplot(countries) +
geom_sf(fill="gray80",color = "black", 
        alpha = 1,linewidth=0.1) + 
geom_point(data = all_data, alpha=1,shape=21,
           aes(x= Longitude, y=Latitude,
           size=Counts, group=Group, colour = Group))+  
coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
         xlim = c(-14, 39), ylim = c(37, 70))+
theme(panel.grid.major = element_blank(), 
      legend.position = ("right"),
         legend.background = element_blank(),
         legend.box.background = element_blank(),
         legend.key = element_blank(), 
         legend.title = element_text(face = "bold",  size = 8),
      axis.text = element_text(size=6, face="bold"),
         axis.title = element_text(face="bold", size = 10),
         plot.title = element_text(face="bold", size = 10),
         legend.text = element_text(face = "bold",
         size = 6), legend.direction="vertical",
         legend.spacing.y = unit(-0, "cm"),
      legend.title.align=0.5) +
theme(panel.background = element_rect(fill = "white"),
        panel.border = element_rect(colour = "black",
        fill=NA, size=1))+
  scale_color_manual(name = "Taxonomic\ngroup", values = c("#009E73", "#D55E00")) +
  scale_size(name = "Species\ncounts") +
  ylab("Latitude") +
xlab("Longitude") +
guides(color = guide_legend(override.aes = list(size=3)))+
scale_x_continuous(breaks = seq(0, 20, by = 10)) +
scale_y_continuous(breaks = seq(40, 60, by = 10))

```
# Results
